---
title: "Coding Challenge6"
author: "Muhtarin Khayer Brohee"
date: "2025-04-04"
output:
  word_document: default
  variant: gfm
  html_document: default
  md_document:
  toc: true
  toc_float: true
---

[Git link](https://github.com/mzb0226/PLPA-6820/tree/main/Coding%20challenge%206)

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Writing Functions and For Loops in R

#### Importance of Writing Your Own Functions and Using Iterations
- Improves **reproducibility** by preventing copy-paste errors  
- Makes code easier to **read, debug, and reuse**  
- Automates **repetitive tasks**, which is crucial when handling large datasets  

---

#### How to Write a Function

In R, a function is a **reusable block of code** that performs a specific task. To write a function, use the syntax:

```r
my_function <- function(input1, input2) {
  # Code to perform some operation
  result <- input1 + input2
  return(result)
}
```
##### Explanation
- The function starts with a name, followed by <- function(...) to define the inputs.

- Code is written inside {} curly brackets.

- The return() statement defines what the function will output.

- I usually write functions at the top of my script or .Rmd file so I can reuse them later.

###  How to Write a For Loop

A `for` loop is used when I want to **repeat an action multiple times**. The basic syntax is:

```r
for (i in 1:5) {
  print(i)
}
```

##### Explanation
- The loop runs the code inside {} for each value of i in the sequence 1:5.

- I write the loop where I need repeated actions, like applying a function to every row in a dataset.

- Loops are useful for automating calculations, simulations, or building data structures.

### Load data

```{r load data, include=TRUE, echo=TRUE,warning=FALSE}
cities <- read.csv("Cities.csv")
```

### Load libraries
library(tidyverse)
```{r load, include=TRUE, warning=FALSE,echo=TRUE}
library(tidyverse)
```

### Haversine distance function
```{r haversine, include=TRUE,warning=FALSE,echo=TRUE}
haversine_dist <- function(lat1, lon1, lat2, lon2) {
  rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a))
  
  earth_radius <- 6378137
  distance_km <- (earth_radius * c)/1000
  
  return(distance_km)
}
```


### Compute distance between Auburn and New York
```{r step5, include=TRUE, echo=TRUE, warning=FALSE}
auburn <- cities %>% filter(city == "Auburn")
nyc <- cities %>% filter(city == "New York")

distance_to_nyc <- haversine_dist(
  lat1 = auburn$lat, lon1 = auburn$long,
  lat2 = nyc$lat, lon2 = nyc$long
)

print(distance_to_nyc)
```


### Loop to calculate distance from all cities to Auburn
```{r loop, include=TRUE, echo=TRUE, warning=FALSE}
#Initialize empty dataframe
distance_df <- data.frame(City1 = character(), City2 = character(), Distance_km = numeric())

# Auburn’s coordinates
auburn_lat <- auburn$lat
auburn_lon <- auburn$long

# Loop through each city except Auburn
for (i in 1:nrow(cities)) {
  if (cities$city[i] != "Auburn") {
    city_name <- cities$city[i]
    city_lat <- cities$lat[i]
    city_lon <- cities$long[i]
    
    dist_km <- haversine_dist(auburn_lat, auburn_lon, city_lat, city_lon)
    
    new_row <- data.frame(City1 = city_name, City2 = "Auburn", Distance_km = dist_km)
    distance_df <- rbind(distance_df, new_row)
  }
}

# View first few distances
head(distance_df)
```

