---
title: "CodingChallenge4"
author: "Muhtarin Khayer Brohee"
date: "2025-02-27"
output:
  md_document:
  variant: gfm
  html_document: default
  pdf_document: default
  word_document: default
toc: true
---
`
## [Git link](https://github.com/mzb0226/PLPA-6820/blob/main/Coding%20Challenge4/Challagne4.md)

## **YAML header**
The YAML (Yet Another Markup Language) header is a metadata block at the beginning of an R Markdown (.Rmd) document. It defines settings for the document, such as the title, author, output format, and table of contents. It is enclosed in triple dashes (---).
Example:
title: "CodingChallenge4"
author: "Muhtarin Khayer Brohee"
date: "2025-02-27"
output:
  md_document:
  variant: gfm
  html_document: default
  pdf_document: default
  word_document: default
toc: true

## **Literate programming**

Literate programming is an approach where code and natural language explanations are combined in a single document. The goal is to create a human-readable and reproducible document that integrates analysis with documentation.

How does R Markdown support literate programming?
 - Code chunks (````{r}```) allow embedding of executable R code.
 - Markdown text enables writing explanations, interpretation, and formatting.
 - Knit output produces reports in HTML, PDF, or Word, making them easy to share.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(ggplot2)
library(markdown)
library(ggplot2)
library(ggpubr)
```
[Article Link](https://pubmed.ncbi.nlm.nih.gov/34587775/)
```{r data_load, include=TRUE, echo=TRUE, warning=FALSE}
DON_data <- read.csv("MycotoxinData.csv", na.strings = "na")
str(DON_data)
```
```{r check_data, include=TRUE,echo=TRUE, warning=FALSE}
# Print column names to check for typos or formatting issues
print(colnames(DON_data))

# Show the first few rows of the dataset
head(DON_data)

# Check the structure of the data
str(DON_data)

```
## **Plot**
```{r data_plot,include=TRUE, echo=TRUE, warning=FALSE}
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


DON_plot <- ggplot(DON_data, aes(x = Treatment, y = DON, fill = Cultivar)) +
  geom_boxplot(outliers = T, outlier.shape = 15) +
  ylab("DON (ppm)") +
  xlab("") +
  geom_jitter(pch = 21, position = position_jitterdodge(), color = "black", alpha = 0.6) + 
  scale_fill_manual(values = c(cbbPalette[[3]], cbbPalette[[4]])) +
  theme_classic() +
  facet_wrap(~Cultivar)
DON_plot

DON_data$Treatment <- factor(DON_data$Treatment, levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70"))

DON_15 <- ggplot(DON_data, aes(x = Treatment, y = X15ADON, fill = Cultivar)) +
  geom_boxplot(outliers = F) +
  ylab("DON (ppm)") +
  xlab("") +
  geom_jitter(pch = 21, position = position_jitterdodge(), color = "black") + 
  scale_fill_manual(values = c(cbbPalette[[3]], cbbPalette[[4]])) +
  theme_classic() +
  facet_wrap(~Cultivar)
DON_15

seedmass <- ggplot(DON_data, aes(x = Treatment, y = MassperSeed_mg, fill = Cultivar)) +
  geom_boxplot(outliers = F) +
  ylab("DON (ppm)") +
  xlab("") +
  geom_jitter(pch = 21, position = position_jitterdodge(), color = "black") + 
  scale_fill_manual(values = c(cbbPalette[[3]], cbbPalette[[4]])) +
  theme_classic() +
  facet_wrap(~Cultivar)
seedmass


```

### **Combined plot**
```{r combined_plot, include=TRUE, echo=TRUE, warning=FALSE}
ggarrange(DON_plot, DON_15, seedmass, labels = "auto", ncol = 3, nrow = 1, common.legend = T)
```


## **Statistical comparison**
```{r stat_plot, include=TRUE, echo=TRUE, warning=FALSE}
stats_donplot <- DON_plot + 
  geom_pwc(aes(group = Treatment), method = "t_test", label = "{p.adj.format}{p.adj.signif}", p.adjust.method = "fdr")

stats_DON_15 <- DON_15 + 
  geom_pwc(aes(group = Treatment), method = "t_test", label = "{p.adj.format}{p.adj.signif}", p.adjust.method = "fdr")

stats_seedmass <- seedmass + 
  geom_pwc(aes(group = Treatment), method = "t_test", label = "{p.adj.format}{p.adj.signif}", p.adjust.method = "fdr")

```
## **File Tree**
```{r file_tree, include=TRUE,echo=TRUE,warning=FALSE}
fs::dir_tree()

```

